How did the introduction of Stemmed Points affect Mobility and Site Occupation during the Late Pleistocene in Korea?

### 2020-12-16 analysis/paper/paper.Rmd
set fig 4 width and make legend text pretty
Fig 5 sizing, label text edit
suppress section numbers in output, stop superflous world map from appearing

```{r}	
quietly <- function(x) {	
  sink(tempfile())	
  on.exit(sink())	
  invisible(force(suppressMessages(x)))	
}	
```
# Introduction

@@

# Background

@@

## Behavioural ecology and forager land use behaviours

@@

# Methods

## Sites and dates

@@

```{r korea-map, fig.cap="Korean Paleolithic sites mentioned in the text.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-draw-the-map.R"))}
quietly(source(here::here("analysis/paper/001-draw-the-map.R")))
knitr::include_graphics(here::here("analysis/figures/001-site-map.png"))

```

## Assemblages

@@

## Statistical methods

@@

# Results

## Artefact volumetric density and retouch frequency

@@

## Composition of toolkit

@@

## Raw materials

@@

## Environmental context

Figure \@ref(fig:site-elevation) shows that people occupied sites located at similar elevations, around 0 to 200 meters above sea level. However, there is a substantial difference in elevation between sites with stemmed points and sites without. The distribution of elevations of sites without stemmed points is generally lower than the elevations of sites with stemmed points. Although not a statistically significant difference in elevations (p = 0.102), this may indicate that forager groups who used stemmed points were able to occupy higher altitudes, while the groups without tended to prefer lower altitudes. Sites without stemmed points reach a maximum elevation of about 150m at around 39ka. Sites containing stemmed points reach a maximum elevation of 175 m at 28ka.

@@

## Demographic context

@@

# Discussion

Our study focussed on two questions: how did the introduction of the new tools change foragers' mobility and landscape use; and how did the new technology change the way people use habitation sites? Overall, our results show that assemblages with stemmed points tend to have higher artefact densities and lower proportions of retouched pieces. Quartz and side scrapers, in addition to cores and choppers, remain dominant in assemblages before and after the introduction of stemmed points. The environmental context of this technological innovation was a gradual decrease in temperature into the LGM.

## Artefact volumetric density and retouch frequency

@@

## Composition of toolkits and raw materials

@@

## Environmental and demographic context

@@

# Conclusion

@@

# Acknowledgements

@@

# References

@@

<div>

</div>

@@

### Colophon

>>>>

```{r}	
quietly <- function(x) {	
  sink(tempfile())	
  on.exit(sink())	
  invisible(force(suppressMessages(x)))	
}	
```
# Introduction {-}

@@

# Background {-}

@@

## Behavioural ecology and forager land use behaviours {-}

@@

# Methods {-}

## Sites and dates {-}

```{r korea-map, fig.cap="Korean Paleolithic sites mentioned in the text.", cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-draw-the-map.R"))}
quietly(source(here::here("analysis/paper/001-draw-the-map.R")))
knitr::include_graphics(here::here("analysis/figures/001-site-map.png"))
```

@@

## Assemblages {-}

@@

## Statistical methods {-}

@@

# Results {-}

## Artefact volumetric density and retouch frequency {-}

@@

## Toolkit composition {-}

@@

## Raw materials {-}

@@

## Environmental context {-}

Figure \@ref(fig:site-elevation) shows that people occupied sites located at similar elevations, around 0 to 200 meters above sea level. However, there is a substantial difference in elevation between sites with stemmed points and sites without. The distribution of elevations of sites without stemmed points is generally lower than the elevations of sites with stemmed points. Although not a statistically significant difference in elevations (p = 0.102), this may indicate that forager groups who used stemmed points were able to occupy higher altitudes, while the groups without tended to prefer lower altitudes. Sites without stemmed points reach a maximum elevation of about 150 m at around 39ka. Sites containing stemmed points reach a maximum elevation of 175 m at 28ka.

@@

## Demographic context {-}

@@

# Discussion {-}

Our study focused on two questions: how did the introduction of the new tools change foragers' mobility and landscape use; and how did the new technology change the way people use habitation sites? Overall, our results show that assemblages with stemmed points tend to have higher artefact densities and lower proportions of retouched pieces. Quartz and side scrapers, in addition to cores and choppers, remain dominant in assemblages before and after the introduction of stemmed points. The environmental context of this technological innovation was a gradual decrease in temperature into the LGM.

## Artefact volumetric density and retouch frequency {-}

@@

## Toolkit composition and raw materials {-}

@@

## Environmental and demographic contexts {-}

@@

# Conclusion {-}

@@

# Acknowledgements {-}

@@

# References {-}

@@

::: {#refs}
:::

@@

### Colophon {-}
